\textsf{Are large language models superhuman chemists?}

### 2024-10-18 src/tex/appendix.tex
add process figure

\begin{figure}[htb]
    \centering
    \includegraphics{figures/flesch_kincaid_reading_ease.pdf}
    \caption{\textbf{Distribution of Flesch-Kincaid reading ease scores of the questions.} The Flesch-Kincaid reading ease score\autocite{flesch1948new} measures how easy a text is to read. It is calculated based on the average number of syllables per word and words per sentence. The higher the score, the easier the text is to read. The distribution of the questions' scores is shown in the histogram. }
    \label{fig:flesch_kincaid_reading_ease}
    \script{wordcloud.py}
\end{figure}

\Cref{fig:question_count_barplot_mcq_vs_general} shows that most questions in our corpus are \gls{mcq}.
A substantial fraction, in contrast to other benchmarks, is open-ended. 
\begin{figure}[htb]
    \centering
    \includegraphics{figures/question_count_barplot_mcq_vs_general.pdf}
    \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of \gls{mcq} and general questions per topic.}
    \label{fig:question_count_barplot_mcq_vs_general}
    \script{plot_statistics.py}
\end{figure}

The corpus of the questions in  \chembench, as shown in \Cref{tab:chembench_corpus}, can be divided according to which topic they belong to or which type of knowledge is required to solve them correctly.

\begin{table*}[!t]
    \centering
        \caption{Summary of \chembench question types.}
    }
    \begin{tabular}{p{2.7cm}cp{10cm}}
        \toprule
        \textbf{Question Type} & \textbf{\%} & \textbf{Example Questions}  \\ 
        \midrule
        Chemical Preference & 35.0 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality). Which of the following two candidates would you prefer for further development? \\
        & & A. [START_SMILES]N#Cc1ccc(OCCCN2CC3CN(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1[END_SMILES] \\
        & & B. [START_SMILES]O=C1CC(c2ccc(CC(NS(=O)(=O)c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S(=O)(=O)N1[END_SMILES] \\
        \midrule
        Technical Chemistry & 1.5 & Which of the following statements is true about the different types of ideal reactors? \\
        & & A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
        & & B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
        & & C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
        & & D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
        \midrule
        Organic Chemistry & 15.5 & What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC[END_RXNSMILES]? \\
        & & A. $E_1$ \\
        & & B. $E_{cb}$ \\
        & & C. $S_N1$ \\
        & & D. $S_N2$ \\
        \midrule
        Toxicity/Safety & 24.2 & What details pertaining to a hazardous substance can be located on a Material Safety Data Sheet (MSDS)? \\
        & & A. All hazards regarding the material \\
        & & B. Physical and chemical properties \\
        & & C. Precautions for safe handling and storage \\
        \midrule
        Analytical Chemistry & 5.8 & Which of the following analytical methods is most appropriate for performing a survey analysis of a solid sample containing various metals? \\
        & & A. X-ray fluorescence analysis \\
        & & B. Differential pulse polarography \\
        & & C. Flame-atomic absorption spectroscopy \\
        & & D. Gas chromatography with flame ionization detector \\
        & & E. Hydride generation atomic absorption spectroscopy \\
        \midrule
        Materials Science & 3.1 & For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact. Why would one choose \\ce{HF} over \\ce{HCl} for this purpose? \\
        & & A. \\ce{F-} forms a stable bonds to the metal ions \\
        & & B. \\ce{HF} has a better water solubility than \\ce{HCl} \\
        & & C. \\ce{HF} has a higher boiling point than \\ce{HCl} \\
        & & D. \\ce{HF} is a weaker acid than \\ce{HCl} \\
        \midrule
        General Chemistry & 5.3 & Which of the following salts is an acidic salt? \\
        & & A. \\ce{NH4Cl} \\
        & & B. \\ce{Na2CO3} \\
        & & C. \\ce{NaH2PO4} \\
        & & D. \\ce{Zn(OH)Cl} \\
        \midrule
        Physical Chemistry & 6.3 & The Born-Oppenheimer (BO) approximation is widely used in computational chemistry, but its accuracy can vary depending on the system. Among the following options, for which system is the Born-Oppenheimer approximation likely to be least applicable? \\
        & & A. \\ce{C60} \\
        & & B. \\ce{CH4} \\
        & & C. \\ce{Fe(CO)5} \\
        & & D. \\ce{H2+} \\
        & & E. \\ce{NaCl} \\
        \midrule
        Inorganic Chemistry & 3.3 & What is the oxidation number of the metal in the compound \\ce{[ZrF7]^{3-}}\\
        \toprule
        \textbf{Reasoning Type} & \textbf{Number of Questions} & \textbf{Example Question} \\
        \midrule
        Knowledge & 770 & Which of the following salts is an acidic salt? \\
        & & A. \\
        \midrule
        Reasoning & 825 & How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
        & & A. By measuring atomic force microscopy \\
        & & B. By measuring fourier transform infrared spectroscopy \\
        & & C. By measuring photoluminescence spectrum \\
        & & D. By measuring the absorption spectroscopy \\
        \midrule
        Intuition & 1001 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality). Which of the following two candidates would you prefer for further development? \\
        & & A. [START_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C@@H]1O[END_SMILES] \\
        & & B. [START_SMILES]Cc1ccccc1N=C(S)N1CCC(NC(=O)c2ccco2)CC1[END_SMILES] \\
        \midrule
        Calculation & 91 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \\pu{595 g mol^{-1}} using end-group analysis, where \\pu{0.1619 g} of PLA was dissolved in \\pu{25 cm^3} of benzyl alcohol and titrated with \\pu{0.0400 mol dm^{-3}} \\ce{KOH} solution. The volume of \\ce{KOH} solution required to reach the endpoint was \\pu{6.81 cm^3}. What is the average number of monomer units in each polymer chain of this sample? \\
        \bottomrule
    \end{tabular}
    \label{tab:chembench_corpus}
\end{table*}

@@

\Cref{fig:parallel_coordinates_overall} shows this data as a parallel coordinates plot. 
This visualization highlights the critical observation that the ranking of the models highly depends on the questions they are evaluated on.
Only very broad benchmarks have the potential to provide a comprehensive view of a model's capabilities. However, even in those cases, the weighting of the different topics is crucial.
Hence, we believe that fine-grained analysis of model performance is vital for the development of future benchmarks.

\begin{figure}[htb]
    \centering
    \includegraphics{figures/parallel_coordinates_overall.pdf}
    \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The parallel coordinates plot shows the performance of the models on the different topics of the \chembench corpus. The performance is measured as the fraction of questions answered completely correctly by the models. }
    \label{fig:parallel_coordinates_overall}
    \script{make_parallel_coordinates_plot.py}
\end{figure}

\begin{figure}[htb]
    \centering
    \includegraphics{figures/parallel_coordinates_tiny.pdf}
    \caption{\textbf{Performance of the models on the different topics of the \enquote{tiny} subset.} The parallel coordinates plot shows the performance of the models on the different topics of the \enquote{tiny} subset. The performance is measured as the fraction of questions answered completely correctly by the models. }
    \label{fig:parallel_coordinates_tiny}
    \script{make_parallel_coordinates_plot.py}
\end{figure}

@@

One might wonder if questions that are more difficult to parse lead to worse performance of the models.
\Cref{fig:reading_ease_vs_model_performance} shows no clear correlation between the reading ease of the questions and the performance of the models.

\begin{figure}
    \centering
    \hspace*{-1cm}
    \includegraphics{figures/reading_ease_vs_model_performance.pdf}
    \caption{\textbf{Model performance as a function of reading ease.} The violin plots show the distribution of reading ease scores for questions answered completely correctly and those not. We do not observe a clear correlation between the reading ease of the questions and the performance of the models. }
    \script{reading_ease_vs_model_performance.py}
    \label{fig:reading_ease_vs_model_performance}
\end{figure}

In addition, we analyzed the performance on questions which requires calculation. For this, we manually labeled questions that require multiple calculation steps. 
We find that the ranking of models is different for questions with our without calculations (\Cref{fig:requires_cal}).

\begin{figure}
    \centering
    \includegraphics{figures/model_overall_cal.pdf}
    \caption{\textbf{Overall model performance for questions with and without calculation steps.} We find that the ranking of models changes if we evaluate them on questions that require and do not require calculations, respectively.}
    \script{requires_calculation_plot.py}
    \label{fig:requires_cal}
\end{figure}

@@

@@

\begin{table}
    \caption{\textbf{Spearman correlation coefficients for the correlation of model performance with molecular features.} The table shows the Spearman correlation coefficient $\rho$ for the correlation of the performance of the models with the number of atoms and the complexity of the molecules. }
    \begin{tabularx}{\textwidth}{XXXX XXXX}
        \toprule
        topic & molecular descriptor & \(\rho\) GPT-4 & \(\rho \) Claude 3 & \( \rho \) Galactica &  \( \rho \) humans  \\ 
        \midrule
        number of \gls{nmr} signals & number of atoms &  \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_number_nmr_peaks.txt}  & \variable{output/correlation_correlation/spearman_num_atoms_human_is_number_nmr_peaks.txt} \\
                              & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_number_nmr_peaks.txt} \\
        \midrule 
        total electron counts & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_electron_counts.txt} \\
        \midrule 
        \gls{smiles} \gls{iupac} name matching & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_name.txt} \\
                & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_name.txt} \\
        \bottomrule
    \end{tabularx}
    \label{tab:correlation_coefficients}
\end{table}

\subsection{Influence of model scale}
To obtain first insights in how the performance \glspl{llm} depends on scale, we tested the \glspl{llm} of the LLaMA series. 
Interestingly, we find that the 7B and 70B models perform comparably, with the 13B showing lower performance (fraction of correct answers for the 7B, 13B, and 70B models are \variable{output/llama/llama_7b.txt}, \variable{output/llama/llama_13b.txt}, \variable{output/llama/llama_70b.txt}).

Note that such analyses are difficult as models are typically not directly comparable in terms of dataset and training protocol.\autocite{biderman2023pythia}

\subsection{Human baseline}
\paragraph{App} To facilitate the collection of responses, we developed a responsive web application in Typescript using the Next.js\autocite{nextjs} app router framework.
This application handles serving the user interface and exposes various \gls{rest} \glspl{api} for relevant operations.
We utilize a MySQL\autocite{mysql} database and Prisma \gls{orm}\autocite{prisma} for efficient database management.
The web application is styled with Tailwind CSS\autocite{tailwindcss} using the shadcn/ui component library and uses NextAuth\autocite{nextauth} for easy and secure user authentication and postMark for sending Emails.
The application is hosted on the Vercel web hosting platform.

@@

We also recorded the time humans took to answer the questions. This time is the time from the question being displayed to the human to the human submitting the answer.
Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (\Cref{fig:human_timing}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\)$\quad$, and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).

\begin{figure}[htb]
    \centering
    \includegraphics{figures/human_timing.pdf}
    \script{analyze_human_data.py}
    \caption{\textbf{Time taken by human scorers to answer questions vs.\ correctness of their answers.} From the plot, it is clear that there is no clear dependence of the correctness of the answers on the time taken by the human scorers to answer the questions.}
    \label{fig:human_timing}
\end{figure}

Additionally, we prompted users to provide additional information about their experience in chemistry. 
While we recorded fine-grained information, e.g., their specialization, we focused on the number of years since the first university-level chemistry course.
\Cref{fig:experience_vs_correctness} shows that the experience of the human scorers was not significantly correlated with the correctness of their answers (\Cref{fig:experience_vs_correctness}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\)$\quad$, and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).

\begin{figure}[htb]
    \centering
    \includegraphics{figures/experience_vs_correctness.pdf}
    \script{analyze_human_data.py}
    \caption{\textbf{Experience of human  scorers vs.\ correctness of their answers.} The experience (in the number of years since the first university-level chemistry course) of the human scorers was not significantly correlated with the correctness of their answers.}
    \label{fig:experience_vs_correctness}
\end{figure}

>>>>

% \begin{figure}[htb]
%     \centering
%     \includegraphics{figures/flesch_kincaid_reading_ease.pdf}
%     \caption{\textbf{Distribution of Flesch-Kincaid reading ease scores of the questions.} The Flesch-Kincaid reading ease score\autocite{flesch1948new} measures how easy a text is to read. It is calculated based on the average number of syllables per word and words per sentence. The higher the score, the easier the text is to read. The distribution of the questions' scores is shown in the histogram. }
%     \label{fig:flesch_kincaid_reading_ease}
%     \script{wordcloud.py}
% \end{figure}

\Cref{fig:question_count_barplot_mcq_vs_general} shows that most questions in our corpus are \gls{mcq}.
A substantial fraction, in contrast to other benchmarks, is open-ended. 
% \begin{figure}[htb]
%     \centering
%     \includegraphics{figures/question_count_barplot_mcq_vs_general.pdf}
%     \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of \gls{mcq} and general questions per topic.}
%     \label{fig:question_count_barplot_mcq_vs_general}
%     \script{plot_statistics.py}
% \end{figure}

@@

@@

% In addition, we analyzed the performance on questions which requires calculation. For this, we manually labeled questions that require multiple calculation steps. 
% We find that the ranking of models is different for questions with our without calculations (\Cref{fig:requires_cal}).

% \begin{figure}
%     \centering
%     \includegraphics{figures/model_overall_cal.pdf}
%     \caption{\textbf{Overall model performance for questions with and without calculation steps.} We find that the ranking of models changes if we evaluate them on questions that require and do not require calculations, respectively.}
%     \script{requires_calculation_plot.py}
%     \label{fig:requires_cal}
% \end{figure}

@@

\subsection{Implementation}

\begin{figure}
    \centering 
    \includegraphics{figures/process.pdf}
    \caption{\textbf{Overview of the benchmarking pipeline implemented in \chembench.} The process begins with JSON files containing task data, which are used to create \texttt{Task} objects and initialize the \texttt{ChemBenchmark}. \texttt{Prompter} objects are then created to handle different types of prompts for instruction-tuned and completion models.
    The \texttt{Task} Iterator differentiates between \gls{mcq} and other question types. For each task type, appropriate prompts are generated and passed to the \gls{llm}. The responses are then processed using regex-based parsing methods specific to \gls{mcq} or numeric answers (after obtaining the relevant part of the response from the instruction-tuned models). 
    The regex-based parsing is elaborate and also able to handle special cases such as scientific notation, or roman numerals.
    If the initial parsing is unsuccessful, the system employs an \gls{llm}-based extraction method as a fallback. The parsed or extracted answers then undergo metric computation and refusal detection.}
    \label{fig:process}
\end{figure}

@@

% \begin{table}
%     \caption{\textbf{Spearman correlation coefficients for the correlation of model performance with molecular features.} The table shows the Spearman correlation coefficient $\rho$ for the correlation of the performance of the models with the number of atoms and the complexity of the molecules. }
%     \begin{tabularx}{\textwidth}{XXXX XXXX}
%         \toprule
%         topic & molecular descriptor & \(\rho\) GPT-4 & \(\rho \) Claude 3 & \( \rho \) Galactica &  \( \rho \) humans  \\ 
%         \midrule
%         number of \gls{nmr} signals & number of atoms &  \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_number_nmr_peaks.txt}  & \variable{output/correlation_correlation/spearman_num_atoms_human_is_number_nmr_peaks.txt} \\
%                               & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_number_nmr_peaks.txt} \\
%         \midrule 
%         total electron counts & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_electron_counts.txt} \\
%         \midrule 
%         \gls{smiles} \gls{iupac} name matching & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_name.txt} \\
%                 & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_name.txt} \\
%         \bottomrule
%     \end{tabularx}
%     \label{tab:correlation_coefficients}
% \end{table}

\subsection{Influence of model scale}
\Cref{fig:model_size_plot} shows the performance of the models as a function of the number of parameters in the model.
We see that the performance of the models correlates with scale for the models of the LLama3 and Llama3.1 herd of models.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/model_size_plot.pdf}
    \caption{\textbf{Performance of models as a function of model size.} The plot shows the performance of the models as a function of the parameter count. The performance is measured as the fraction of questions answered correctly by the models. We see that the performance of the models correlates with scale for the models of the LLama3 and Llama3.1 herd of models.}
    \script{performance_vs_model_size.py}
    \label{fig:model_size_plot}
\end{figure}

\subsection{Human baseline}
\paragraph{App} To facilitate the collection of responses, we developed a responsive web application in Typescript using the Next.js\autocite{nextjs} app router framework.
This application handles serving the user interface and exposes various \gls{rest} \glspl{api} for relevant operations.
We utilize a Postgresql.
The web application is styled with Tailwind CSS\autocite{tailwindcss} using the shadcn/ui component library and uses NextAuth\autocite{nextauth} for easy and secure user authentication.
The application is hosted on the Vercel web hosting platform.

In the applications, human participants were presented with molecules as rendered drawings and SMILES strings. \LaTeX\xspace equations and chemical equations were rendered using MathJax\autocite{mathjax}.

@@

We also recorded the time humans took to answer the questions. This time is the time from the question being displayed to the human to the human submitting the answer.
%Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (\Cref{fig:human_timing}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\)$\quad$, and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).

\begin{figure}[htb]
    \centering
    \includegraphics{figures/human_timing.pdf}
    \script{analyze_human_data.py}
    \caption{\textbf{Time taken by human scorers to answer questions vs.\ correctness of their answers.} From the plot, it is clear that there is no clear dependence of the correctness of the answers on the time taken by the human scorers to answer the questions. However, we see that human scorers typically took longer to correctly answer questions with tool use.}
    \label{fig:human_timing}
\end{figure}

Additionally, we prompted users to provide additional information about their experience in chemistry. 
While we recorded fine-grained information, e.g., their specialization, we focused on the number of years since the first university-level chemistry course.
\Cref{fig:experience_vs_correctness} shows that the experience of the human scorers was correlated with the correctness of their answers (\Cref{fig:experience_vs_correctness}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).

\begin{figure}[htb]
    \centering
    \includegraphics{figures/experience_vs_correctness.pdf}
    \script{analyze_human_data.py}
    \caption{\textbf{Experience of human  scorers vs.\ correctness of their answers.} The experience (in the number of years since the first university-level chemistry course) of the human scorers wasp correlated with the correctness of their answers.}
    \label{fig:experience_vs_correctness}
\end{figure}
